---
title: "p8105_hw6_jp3665"
author: "Joana Petrescu"
date: "12/8/2020"
output: github_document
---

```{r message=FALSE, warning=FALSE}

library(tidyverse)
## install.packages("janitor")
library(janitor)
## install.packages("modelr")
library(modelr)

```


## Washington Post homicide data

This data gathered by the Washington Post contains information about the victim, location, and disposition of over 52,000 homicides that occurred over the past decade in 50 of the largest cities in the U.S.

Import crime data, create a `city_state` variable, filter out cities that do not report victim race, add logical column `solved` representing whether the case has been solved (Closed with arrest or Closed without arrest), filter to include only crimes with victims that were `White` or `Black`.

```{r warning=FALSE}

homicide_data <- read.csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv")

homicide_data <- homicide_data %>%
  mutate(.keep = "all", city_state = paste(city, state, sep = ", ")) %>%
  filter(city_state != c("Phoenix, AZ", "Dallas, TX", "Kansas City, MO", "Tulsa, AL")) %>%
  mutate(solved = as.numeric(disposition != c("Open/No arrest"))) %>%
  mutate(victim_age = as.numeric(victim_age)) %>%
  filter(victim_race %in% c("White", "Black")) %>%
  select(city_state, solved, victim_age, victim_race, victim_sex)



head(homicide_data)

```

### Generalized linear model for solving homicides by race

Using Baltimore, MD as an example, use a generalized linear model to calculate odds ratio and confidence intervals for solving homicides comparing non-white victims to white victims, keeping age and sex fixed.

```{r warning=FALSE}

baltimore_glm <- homicide_data %>%
  filter(city_state == "Baltimore, MD") %>%
  mutate(victim_race = factor(victim_race, levels = c("White", "Black"))) %>%
  glm(solved ~ victim_age + victim_race + victim_sex, data = ., family = binomial()) %>%
  broom::tidy() %>%
  mutate(OR = exp(estimate), OR_min = exp(estimate - std.error), OR_max = exp(estimate + std.error)) %>%
  mutate(across(is.numeric, ~ round(., 3))) %>%
  mutate(CI = (paste(OR_min, OR_max, sep = "-"))) %>%
  select(term, OR, CI)

baltimore_glm %>% filter(term == "victim_raceBlack") %>% knitr::kable()


```

Iterate across all cities to calculate odds ratio and confidence intervals for solving homicides comparing non-white victims to white victims, keeping age and sex fixed.

```{r warning=FALSE}

homicide_data <- homicide_data %>%
  mutate(victim_race = factor(victim_race, levels = c("White", "Black")))

homicide_nest <- homicide_data %>%
  nest(data = -city_state) %>%
  filter(city_state != "Tulsa, AL")

homicide_glm <- homicide_nest %>%
  mutate(
    results = map(data, ~glm(solved ~ victim_age + victim_race + victim_sex, data = .x, family = binomial()) %>%
                    broom::tidy())
  ) %>%
  select(-data) %>%
  unnest(results) %>%
  mutate(OR = exp(estimate), OR_min = exp(estimate - std.error), OR_max = exp(estimate + std.error)) %>%
  mutate(across(is.numeric, ~ round(., 3))) %>%
  mutate(CI = (paste(OR_min, OR_max, sep = "-"))) %>%
  filter(term == "victim_raceBlack") 

homicide_glm %>%  select(city_state, OR, CI) %>% knitr::kable()


```
Plot estimated odds ratios and confidence intervals for solving homicides comparing non-white victims to white victims, keeping age and sex fixed.

```{r}

homicide_glm %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>%
  ggplot(aes(x = city_state, y = OR)) +
  geom_errorbar(aes(ymin = OR_min, ymax = OR_max)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90)) +
  xlab("City, State") +
  ylab("Estimated Odds Ratio") +
  labs(title = "Estimated Odds Ratio for Solved Homicides Comparing Black to White Victims")

```


## Factors affecting children's birthweight

Load birthweight data and recode variables

```{r}

birthweight_data <- read.csv("./data/birthweight.csv") %>% 
  clean_names() %>%
  mutate(babysex = recode(babysex, "1" = "male", "2" = "female")) %>%
  mutate(frace = recode(frace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other", "9" = "Unknown")) %>%
  mutate(malform = recode(malform, "0" = "absent", "1" = "present")) %>%
  mutate(mrace = recode(mrace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other"))
  

```

### Linear model for factors affecting children's birthweight

Exploring correlations of mother's weight gain, gestational age, pre-pregnancy BMI, and pre-pregnancy weight with birth weight. Mother's weight gain and gestational age seem to be positively correlated with birth weight in this dataset whereas there seems to be no association between pre-pregnancy BMI or pre-pregnancy weight with birth wight.

```{r}

birthweight_data %>% ggplot(aes(x = wtgain, y = bwt)) +
  geom_point()

birthweight_data %>% ggplot(aes(x = gaweeks, y = bwt)) +
  geom_point()

birthweight_data %>% ggplot(aes(x = ppbmi, y = bwt)) +
  geom_point()

birthweight_data %>% ggplot(aes(x = ppwt, y = bwt)) +
  geom_point()

```

In addition to the positive correlations identified above, I know that research has shown that smoking during pregnancy is associated with decreased birth weight. Therefore, my proposed linear model for this birth weight data uses gestational age, mother's weight gain during pregnancy, and average number of cigarettes smoked per day during pregnancy. The second and third linear models are what I will compare my proposed model to.

```{r}

smoking_fit <- lm(bwt ~ gaweeks + smoken + wtgain, data = birthweight_data) 

length_fit <- lm(bwt ~ blength + gaweeks, data = birthweight_data)

circumference_fit <- lm(bwt ~ (bhead + blength + babysex)^2, data = birthweight_data)

```


Plot model birth weight predictions and residuals against birth weight.


```{r}

smoking_fit %>% broom::tidy() %>% select(term, estimate, p.value) %>% knitr::kable(digits = 200)

birthweight_data <- birthweight_data %>% modelr::add_predictions(smoking_fit) %>% modelr::add_residuals(smoking_fit)

birthweight_data %>% ggplot(aes(x = bwt, y = pred)) +
  geom_point()

birthweight_data %>% ggplot(aes(x = bwt, y = resid)) +
  geom_point()

```
These plots demonstrate that my proposed model is insufficient to explain variance in birth weight. The residuals for my model are close to zero when the birth weight is ~3000g but gets up to 1000-2000g when birth weight is low (<2500g) or high (>3500 g). This means that my model can predict birth weights ranging from 2500-3500g relatively accurately but cannot do so outside of this range.

## Crossvalidation of candidate linear models

```{r}

cv_df <- crossv_mc(birthweight_data, 100)

cv_df <- cv_df %>%
  mutate(
    smoking_mod = map(train, ~lm(bwt ~ gaweeks + smoken + wtgain, data = .x)),
    length_mod = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    circumference_mod = map(train, ~lm(bwt ~ (bhead + blength + babysex)^2, data = .x))
  ) %>%
  mutate(
    rmse_smoking = map2_dbl(smoking_mod, test, ~rmse(model = .x, data =.y)),
    rmse_length = map2_dbl(length_mod, test, ~rmse(model = .x, data = .y)),
    rmse_circumference = map2_dbl(circumference_mod, test, ~rmse(model = .x, data = .y))
  )

cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()

```
Based on these calculated RMSE values, the model using head circumference, length, sex, and all interactions between these variables is the most accurate representation of this birth weight data, while the linear model that I proposed which only takes into account gestational gestational age, mother's weight gain during pregnancy and average number of cigarettes smoked per day during pregnancy performed the worst.

## Central Park weather data

Load and clean data

```{r}

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```

